Delayed effects of [responses to?] climate lead to demographic divergence in a fragmented Amazonian landscape

### 2021-03-15 analysis/paper/paper.Rmd

\
1. Understanding how landscape structure and environmental conditions act to influence population dynamics is fundamental to many of the emerging conceptual frameworks for studying and conserving fragmented landscapes [@didhamRethinkingConceptualFoundations2012; @driscollConceptualDomainMatrix2013].

@@

      * This study adds to the growing body of work that finds delayed effects of climate drivers on vital rates without *a priori* assumptions about the influence of particular climate windows [@tenhumbergTimelaggedEffectsWeather2018; @tellerLinkingDemographyDrivers2016; @eversLaggedDormantSeason2021]. 
      * We found evidence for delayed effects of drought which are sometimes better predictors of vital rates than more recent conditions. 
      * Delayed effects of climate on plant vital rates may be the rule, rather than the exception. 
      * distributed lag non-linear models offer an unbiased way of identifying these delayed effects (but see @pierreCriticalWindowsMethod2020 for an alternative method). <!--# maybe the previous sentence goes in methods? -->

@@

- I gather that in a good year, plants can store carbon and nutrients in rhizomes and then mobilize as needed and the opposite is true in a bad year. Flowering is a big investment so perhaps flowering in one year means that you don’t flower the following one. If that is the case, it would make sense to include growth in the previous year as a covariate in all analyses rather than do this indirectly via a potential effect of past SPEI on past growth. It will also be easier to explain as you write the paper. **Emilio will look into physiological responses to light and H20 in the horticultural and ecological literature, eg. Westerband papers, esp if focusing on herbs** Basically anything on how more or less precip influences vital rates or how plants respond to them, esp if linked to changes in canopy openness, litterfall, flowering in ENSO years, Bruna transplant experiments, Heliconia germination in secondary forest, important to emphasize: results are often counter-intuitive.

>>>>

1.  Understanding how landscape structure and environmental conditions act to influence population dynamics is fundamental to many of the emerging conceptual frameworks for studying and conserving fragmented landscapes [@didhamRethinkingConceptualFoundations2012; @driscollConceptualDomainMatrix2013].

@@

    -   This study adds to the growing body of work that finds delayed effects of climate drivers on vital rates without *a priori* assumptions about the influence of particular climate windows [@tenhumbergTimelaggedEffectsWeather2018; @tellerLinkingDemographyDrivers2016; @eversLaggedDormantSeason2021].
    -   We found evidence for delayed effects of drought which are sometimes better predictors of vital rates than more recent conditions.
    -   Delayed effects of climate on plant vital rates may be the rule, rather than the exception.
    -   distributed lag non-linear models offer an unbiased way of identifying these delayed effects (but see @pierreCriticalWindowsMethod2020 for an alternative method). <!--# maybe the previous sentence goes in methods? -->

@@

-   I gather that in a good year, plants can store carbon and nutrients in rhizomes and then mobilize as needed and the opposite is true in a bad year.
    Flowering is a big investment so perhaps flowering in one year means that you don't flower the following one.
    If that is the case, it would make sense to include growth in the previous year as a covariate in all analyses rather than do this indirectly via a potential effect of past SPEI on past growth.
    It will also be easier to explain as you write the paper.
    **Emilio will look into physiological responses to light and H20 in the horticultural and ecological literature, eg. Westerband papers, esp if focusing on herbs** Basically anything on how more or less precip influences vital rates or how plants respond to them, esp if linked to changes in canopy openness, litterfall, flowering in ENSO years, Bruna transplant experiments, Heliconia germination in secondary forest, important to emphasize: results are often counter-intuitive.

-   cite Esteban et al. 2021 for negative effects of large positive SPEI

-   The increase in flowering rate of plants following a drought could be a bet-hedging strategy in response to stress [@nihadRelationshipStressFlowering2018] <!--# add another more general citation. e.g. S. Joe Wright trees flowering after ENSO -->.
